Modeling the Linux page cache for accurate simulation of data-intensive applications

### 2021-05-11 paper_io_sim.tex

    \begin{algorithm}\caption{File chunk read simulation in I/O Controller}
    \label{alg:read}
        \small
        \begin{algorithmic}[1]
            \Input
                \Desc{cs}{chunk size}
                \Desc{fn}{file name}
                \Desc{fs}{file size (assumed to fit in memory)}
                \Desc{mm}{MemoryManager object}
                \Desc{sm}{storage simulation model}
               \EndInput
               \State disk\_read = min(cs, fs - mm.cached(fn)) \Comment{To be read from disk}
               \State cache\_read = cs - disk\_read \Comment{To be read from cache}
               \State required\_mem = cs + disk\_read
               \State mm.flush(required\_mem - mm.free\_mem - mm.evictable, fn)
               \State mm.evict(required\_mem - mm.free\_mem, fn)
               \If {disk\_read $>$ 0}  \Comment{Read uncached data}
               \State sm.read(disk\_read)
               \State mm.add\_to\_cache(disk\_read, fn)
               \EndIf
               \If {cache\_read $>$ 0} \Comment{Read cached}
               \State mm.cache\_read(cache\_read)
            \EndIf
            \State mm.use\_anonymous\_mem(cs)
        \end{algorithmic}
    \end{algorithm}
    As mentioned previously, our model reads and writes file chunks in a
    round-robin fashion. To read a file chunk, simulated applications send
    chunk read requests to the I/O Controller which processes them using
    Algorithm~\ref{alg:read}. First, we calculate the amount of uncached
    data that needs to be read from disk, and the remaining amount is read
    from cache (line 7-8). The amount of memory required to read the chunk
    is calculated, corresponding to a copy of the chunk in anonymous memory
    and a copy of the chunk in cache (line 9).
    If there is not enough available memory, the Memory Manager is called
    to flush dirty data (line 10). If necessary, flushing is complemented by
    eviction (line 11). Note that, when called with negative arguments, functions
    \texttt{flush} and \texttt{evict} simply return and do not do anything. Then,
    if the block requires
    uncached data, the memory manager is called to read data from disk and to add this
    data to cache (line 14).
    If cached data needs to be read, the Memory Manager is called to simulate
    a cache read  and update the corresponding data blocks accordingly (line 17).
    Finally, the memory manager is called to deallocate the amount of anonymous memory used by the application (line 19).

    \begin{algorithm}\caption{File chunk write simulation in I/O Controller}
    \label{alg:write}
        \small
        \begin{algorithmic}[1]
            \Input
                \Desc{cs}{chunk size}
                \Desc{fn}{file name}
                \Desc{mm}{MemoryManager object}
               \EndInput
            \State remain\_dirty = dirty\_ratio * mm.avail\_mem - mm.dirty
            \If {remain\_dirty $>$ 0} \Comment{Write to memory}
                \State mm.evict(min(cs, remain\_dirty) - mm.free\_mem)
                \State mem\_amt = min(cs, mm.free\_mem)
                \State mm.write\_to\_cache(fn, mem\_amt)
            \EndIf
            \State remaining = cs - mem\_amt
            \While {remaining $>$ 0}  \Comment{Flush to disk, then write to cache}
                \State mm.flush(cs - mem\_amt)
                \State mm.evict(cs - mem\_amt  - mm.free\_mem)
                \State to\_cache = min(remaining, mm.free\_mem)
                \State mm.write\_to\_cache(fn, to\_cache)
                \State remaining = remaining - to\_cache
            \EndWhile

>>>>

    \begin{algorithm}\caption{File chunk read simulation in I/O Controller}
    \label{alg:read}
        \small
        \begin{algorithmic}[1]
        \Input
            \Desc{cs}{chunk size}
            \Desc{fn}{file name}
            \Desc{fs}{file size (assumed to fit in memory)}
            \Desc{mm}{MemoryManager object}
            \Desc{sm}{storage simulation model}
           \EndInput
           \State disk\_read = min(cs, fs - mm.cached(fn)) \Comment{To be read from disk}
           \State cache\_read = cs - disk\_read \Comment{To be read from cache}
           \State required\_mem = cs + disk\_read
           \State mm.flush(required\_mem - mm.free\_mem - mm.evictable)
           \State mm.evict(required\_mem - mm.free\_mem)
           \If {disk\_read $>$ 0}  \Comment{Read uncached data}
           \State sm.read(disk\_read)
           \State mm.add\_to\_cache(disk\_read, fn)
           \EndIf
           \If {cache\_read $>$ 0} \Comment{Read cached}
           \State mm.cache\_read(cache\_read)
        \EndIf
        \State mm.use\_anonymous\_mem(cs)
        \end{algorithmic}
    \end{algorithm}
    As mentioned previously, our model reads and writes file chunks in a
    round-robin fashion. To read a file chunk, simulated applications send
    chunk read requests to the I/O Controller which processes them using
    Algorithm~\ref{alg:read}. First, we calculate the amount of uncached
    data that needs to be read from disk, and the remaining amount is read
    from cache (line 7-8). The amount of memory required to read the chunk
    is calculated, corresponding to a copy of the chunk in anonymous memory
    and a copy of the chunk in cache (line 9).
    If there is not enough available memory, the Memory Manager is called
    to flush dirty data (line 10). If necessary, flushing is complemented by
    eviction (line 11). Note that, when called with negative arguments, functions
    \texttt{flush} and \texttt{evict} simply return and do not do anything. Then,
    if the block requires
    uncached data, the memory manager is called to read data from disk and to add this
    data to cache (line 14).
    If cached data needs to be read, the Memory Manager is called to simulate
    a cache read  and update the corresponding data blocks accordingly (line 17).
    Finally, the memory manager is called to deallocate the amount of anonymous memory used by the application (line 19).

    \begin{algorithm}\caption{File chunk write simulation in I/O Controller}
    \label{alg:write}
        \small
        \begin{algorithmic}[1]
        \Input
            \Desc{cs}{chunk size}
            \Desc{fn}{file name}
            \Desc{mm}{MemoryManager object}
           \EndInput
        \State remain\_dirty = dirty\_ratio * mm.avail\_mem - mm.dirty
        \If {remain\_dirty $>$ 0} \Comment{Write to memory}
            \State mm.evict(min(cs, remain\_dirty) - mm.free\_mem)
            \State mem\_amt = min(cs, mm.free\_mem)
            \State mm.write\_to\_cache( mem\_amt, fn)
        \EndIf
        \State remaining = cs - mem\_amt
        \While {remaining $>$ 0}  \Comment{Flush to disk, then write to cache}
            \State mm.flush(cs - mem\_amt)
            \State mm.evict(cs - mem\_amt  - mm.free\_mem)
            \State to\_cache = min(remaining, mm.free\_mem)
            \State mm.write\_to\_cache(to\_cache, fn)
            \State remaining = remaining - to\_cache
        \EndWhile
